Tight Generalization Bounds for Hard Parameter Sharing in High-dimensional Linear Regression

### 2020-10-19 arxiv/proof_examples.tex
finish section D

For the bias term, since the spectrum of $(X^{(1)})^{\top} X^{(1)}$ is tightly concentrated by Fact \ref{fact_minv}, we have that
\begin{align}
 \frac{(\sqrt{n_1}-\sqrt{p})^4 \cdot (1- p^{-c_\varphi})}{p} \bigtr{\hat \Sigma^{-2}}  &\le {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] \label{eq_bias111}\\
 &\le \frac{(\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi})}{p} \bigtr{\hat\Sigma^{-2}}.\nonumber
\end{align}
%To obtain this inequality, we used
%$$ (\sqrt{n_1}-\sqrt{p})^4 \cdot (1-p^{-c_\varphi}) \preceq \left((X^{(1)})^\top X^{(1)} \right)^2 \preceq (\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi}) \quad \text{with high probability},$$
%by Fact \ref{fact_minv} (ii), and the fact that for the product of two PSD matrices, its trace is always nonnegative.
Using the bias limit \eqref{lem_cov_derv_eq} with $\Sigma^{(1)}=\Sigma^{(2)}=\Lambda=V=\id_{p\times p}$, and $w = e_i$ (the $i$-th coordinate vector), we have w.h.p.
\begin{align*}%\label{eq_bias112}
 \left(\hat\Sigma^{-2}\right)_{ii} = \frac{1}{(n_1+n_2)^2} \left[\frac{a_3+a_4+1}{(a_1+a_2)^2} +\OO(p^{-c_\varphi})\right].
\end{align*}
We solve the self-consistent equations \eqref{eq_a34extra} given $a_1, a_2$, and obtain
	\begin{align*}
		a_3 = \frac{p\cdot n_1}{(n_1 +n_2)(n_1 +n_2 - p)}, \quad
		&  a_4 = \frac{p\cdot n_2}{(n_1 + n_2)(n_1 + n_2 - p)}. %\label{simplesovlea34}
	\end{align*}
Applying them to the above equation, we obtain that w.h.p.
\begin{align*}%\label{eq_bias113}
 \left(\hat\Sigma^{-2}\right)_{ii} =  \frac{1}{(n_1+n_2)^2} \left[  \frac{ (n_1+n_2)^3 }{(n_1+n_2-p)^3} +\OO(p^{-c_\varphi})\right].
\end{align*}
In fact, we need this estimate to hold simultaneously for all $1\le i \le p$ with high probability. For this purpose, we shall use equation \eqref{apply derivlocal} to get that
\begin{align*}%\label{eq_bias113}
\left| \left(\hat\Sigma^{-2}\right)_{ii} - \frac{1}{(n_1+n_2)^2}\cdot \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\prec \frac{p^{-\frac{\varphi-4}{2\varphi}}}{(n_1+n_2)^2}
\end{align*}
on a high probability event that does not depend on $i$. Then using Fact \ref{lem_stodomin} (i), we obtain that 
\begin{align*} 
\left| \sum_{i=1}^p \left(\hat\Sigma^{-2}\right)_{ii} - \frac{p}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\prec \frac{p^{1-\frac{\varphi-4}{2\varphi}}}{(n_1+n_2)^2}
\end{align*}
with high probability, which by Definition \ref{stoch_domination} gives that
\begin{align}\label{eq_bias113}
\left| p^{-1}\sum_{i=1}^p \left(\hat\Sigma^{-2}\right)_{ii} - \frac{1}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\le \frac{p^{-c_\varphi}}{(n_1+n_2)^2}
\end{align}
with high probability. Inserting equation \eqref{eq_bias113} into equation \eqref{eq_bias111}, we get that
\begin{align*}
		& \frac{(\sqrt{n_1}-\sqrt{p})^4 \cdot (1- p^{-c_\varphi}) - n_1^2 \cdot (1+\OO(p^{-c_\varphi}))}{(n_1+n_2)^2}\cdot   \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\nonumber \\
		\le &  {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] - \frac{n_1^2}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3} \nonumber\\
		\le &  \frac{(\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi}) - n_1^2 \cdot (1+\OO(p^{-c_\varphi}))}{(n_1+n_2)^2}\cdot   \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}
\end{align*}
with high probability.

Combining equation \eqref{eq_var112} and the above with Claim \ref{claim_reduction}, we get that
\begin{align*}
	\left|L(\hat{\beta}_2^{\MTL})  - \frac{\sigma^2 p}{n_1+n_2-p} -  \frac{2d^2 n_1^2 (n_1+n_2)}{(n_1 + n_2 - p)^3}\right| \le   \left[\left( 1+\sqrt{\frac{p}{n_1}}\right)^4-1\right] \cdot \frac{2d^2n_1^2 (n_1+n_2)}{(n_1 + n_2 - p)^3} \nonumber\\
+C\left( p^{-c_\varphi} (\sigma^2 + d^2)+ \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) \right).
\end{align*}
Then using the conditions $\sigma^2 \lesssim  \kappa^2$ and $d^2 \le p^{-\e}{\kappa^2}$, we conclude the proof.
\end{proof}

>>>>

For the bias term, since the spectrum of $(X^{(1)})^{\top} X^{(1)}$ is tightly concentrated by Fact \ref{fact_minv}, we have that
\begin{align}
 \frac{(\sqrt{n_1}-\sqrt{p})^4 \cdot (1- p^{-c_\varphi})}{p} \bigtr{\hat \Sigma^{-2}}  &\le {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] \label{eq_bias111}\\
 &\le \frac{(\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi})}{p} \bigtr{\hat\Sigma^{-2}}.\nonumber
\end{align}
%To obtain this inequality, we used
%$$ (\sqrt{n_1}-\sqrt{p})^4 \cdot (1-p^{-c_\varphi}) \preceq \left((X^{(1)})^\top X^{(1)} \right)^2 \preceq (\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi}) \quad \text{with high probability},$$
%by Fact \ref{fact_minv} (ii), and the fact that for the product of two PSD matrices, its trace is always nonnegative.
Using the bias limit \eqref{lem_cov_derv_eq} with $\Sigma^{(1)}=\Sigma^{(2)}=\Lambda=V=\id_{p\times p}$, and $w = e_i$ (the $i$-th coordinate vector), we have w.h.p. (via a union bound)
\begin{align*}%\label{eq_bias112}
 e_i^{\top}\hat\Sigma^{-2} e_i = \frac{1}{(n_1+n_2)^2} \left[\frac{a_3+a_4+1}{(a_1+a_2)^2} +\OO(p^{-c_\varphi})\right], \text{ for all } i = 1, 2, \dots, p.
\end{align*}
We solve the self-consistent equations \eqref{eq_a34extra} given $a_1, a_2$, and obtain
\begin{align*}
		a_3 = \frac{p\cdot n_1}{(n_1 +n_2)(n_1 +n_2 - p)}, \quad
		&  a_4 = \frac{p\cdot n_2}{(n_1 + n_2)(n_1 + n_2 - p)}. %\label{simplesovlea34}
\end{align*}
Applying them to the above equation, we obtain
\begin{align*}%\label{eq_bias113}
 e_i^{\top}\hat\Sigma^{-2}e_i =  \frac{1}{(n_1+n_2)^2} \left[  \frac{ (n_1+n_2)^3 }{(n_1+n_2-p)^3} +\OO(p^{-c_\varphi})\right].
\end{align*}
%In fact, we need this estimate to hold simultaneously for all $1\le i \le p$ with high probability. For this purpose, we shall use equation \eqref{apply derivlocal} to get that
%\begin{align*}%\label{eq_bias113}
%\left| \left(\hat\Sigma^{-2}\right)_{ii} - \frac{1}{(n_1+n_2)^2}\cdot \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\prec \frac{p^{-\frac{\varphi-4}{2\varphi}}}{(n_1+n_2)^2}
%\end{align*}
%on a high probability event that does not depend on $i$. Then using Fact \ref{lem_stodomin} (i), we obtain that
%\begin{align*}
%\left| \sum_{i=1}^p \left(\hat\Sigma^{-2}\right)_{ii} - \frac{p}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\prec \frac{p^{1-\frac{\varphi-4}{2\varphi}}}{(n_1+n_2)^2}
%\end{align*}
%with high probability, which by Definition \ref{stoch_domination} gives that
%\begin{align}\label{eq_bias113}
%\left| p^{-1}\sum_{i=1}^p \left(\hat\Sigma^{-2}\right)_{ii} - \frac{1}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\right|\le \frac{p^{-c_\varphi}}{(n_1+n_2)^2}
%\end{align}
%with high probability.
Applying the above equation to equation \eqref{eq_bias111}, we get the desired result for the bias term.
%\begin{align*}
%		& \frac{(\sqrt{n_1}-\sqrt{p})^4 \cdot (1- p^{-c_\varphi}) - n_1^2 \cdot (1+\OO(p^{-c_\varphi}))}{(n_1+n_2)^2}\cdot   \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}\nonumber \\
%		\le &  {p}^{-1}\tr\left[\hat\Sigma^{-2} \left((X^{(1)})^\top X^{(1)} \right)^2\right] - \frac{n_1^2}{(n_1+n_2)^2}\cdot  \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3} \nonumber\\
%		\le &  \frac{(\sqrt{n_1}+\sqrt{p})^4 \cdot (1+ p^{-c_\varphi}) - n_1^2 \cdot (1+\OO(p^{-c_\varphi}))}{(n_1+n_2)^2}\cdot   \frac{(n_1+n_2)^3}{(n_1+n_2-p)^3}.
%\end{align*}
Combining the bias and variance estimates, we get that
\begin{align*}
	\left|L(\hat{\beta}_2^{\MTL})  - \frac{\sigma^2 p}{n_1+n_2-p} -  \frac{2d^2 n_1^2 (n_1+n_2)}{(n_1 + n_2 - p)^3}\right| \le   \left[\left( 1+\sqrt{\frac{p}{n_1}}\right)^4-1\right] \cdot \frac{2d^2n_1^2 (n_1+n_2)}{(n_1 + n_2 - p)^3} \nonumber\\
+C\left( p^{-c_\varphi} (\sigma^2 + d^2)+ \frac{d^4 + \sigma^2 d^2}{\kappa^2}+p^{-1/2+2c}\kappa^2 +  p^{-1/4+c} (\sigma^2 +d^2) \right).
\end{align*}
Since $\sigma^2 \lesssim  \kappa^2$ and $d^2 \le p^{-\e}{\kappa^2}$ by our assumption, we obtain the desired result.
The proof is complete.
\end{proof}

### 2020-10-19 arxiv/proof_gaussian_case.tex
finish section D

\begin{theorem}\label{thm_main_RMT_app}
	In the setting of Theorem \ref{thm_main_RMT}, the bias equation \eqref{eq_bias_2task} satisfies the following limit with high probability for any fixed unit vector $w\in\real^p$:
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - \frac{1}{(n_1+n_2)^2}{\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \le  \frac{p^{-c_{\varphi}}}{(n_1+n_2)^2},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \ \
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
%For the rest of this section, we present the proof of Theorem \ref{thm_main_RMT} and Theorem \ref{thm_main_RMT_app}

>>>>

\begin{theorem}\label{thm_main_RMT_app}
	Let $S$ be an arbitrary subset of the unit sphere in dimension $p$ whose size is polynomial in $p$.
	In the setting of Theorem \ref{thm_main_RMT}, the bias equation \eqref{eq_bias_2task} satisfies the following limit with high probability for any unit vector $w\in S$:
			\begin{align}\label{lem_cov_derv_eq}
				\bigabs{w^{\top} \Sigma^{(1)} \bigbrace{\hat{\Sigma}^{-1} \Sigma^{(2)} \hat{\Sigma}^{-1} - \frac{1}{(n_1+n_2)^2}{\Sigma^{(2)}}^{-1/2} V {\frac{a_3 \Lambda^2 + (a_4 + 1)\id}{(a_1 \Lambda^2 + a_2\id)^2}} V^{\top} {\Sigma^{(2)}}^{-1/2}} \Sigma^{(1)} w} \le  \frac{p^{-c_{\varphi}}}{(n_1+n_2)^2},
			\end{align}
				where $a_{3}$ and $a_4$ are the solutions of the following self-consistent equations % with $b_k = \frac1{p}\sum_{i=1}^p \frac{\lambda_i^{2k}} {(\lambda_i^2 a_1 + a_2)^2}$, for $k = 0, 1, 2$:
			\begin{align}\label{eq_a34extra}
				a_3 + a_4 = \frac{1}{n_1 + n_2}\sum_{i=1}^p \frac{1}{\lambda_i^2 a_1 + a_2}, \ \
				a_3 + \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 (a_2 a_3-a_1 a_4 )}{(\lambda_i^2 a_1 + a_2)^2} = \frac{1}{n_1 + n_2} \sum_{i=1}^p \frac{\lambda_i^2 a_1}{(\lambda_i^2 a_1 + a_2)^{2}}.
%				\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
%				= b_0.
			\end{align}
\end{theorem}
%For the rest of this section, we present the proof of Theorem \ref{thm_main_RMT} and Theorem \ref{thm_main_RMT_app}
